The Talking Heads Experiment

### 2015-04-22 chap5/chap5.tex
fixed some more widows/orphans
hyphenation, punctuation, positioning, etc for chapters 1-6

The semiotic square captures the four entities in 
a linguistic interaction: 
the {\itshape referent}, which is the object in the physical world that the 
speaker wants to communicate, the {\itshape segmented image}, which is 
the internal perception of the referent, the {\itshape meaning}, which is a 
category or combination of categories that picks out the 
referent in the present context, and 
the {\itshape utterance}, which is the word form 
or set of word forms transmitted by the speaker. In the previous 
chapters we looked at two sides of the semiotic
square. The relation between the real world and the
perceived image was studied in chapter 3 and the relation
between the perceived image and a conceptualisation that 
could act as the meaning for a language communication was 
studied in chapter 4. We now turn to the next side of 
the semiotic square: the relation between meaning and 
utterance.

@@

Immediately we observe a major difficulty. In realistic
language games, where agents cannot inspect each others' 
brain states nor transmit meanings directly, 
there is no feedback about the 
meaning of a word, only about the referent. For example, 
when a speaker says `wabo' and the hearer has correctly pointed 
to the referent that the speaker intended, neither the 
speaker nor the hearer can know whether they were 
using the same meaning. They can only know that they 
arrived at the same referent. When a game fails, the speaker 
can only point to the topic and the hearer then 
tries to figure out what possible meaning 
could have been applicable. The speaker cannot
communicate directly the ``right'' meaning, and very often more 
than one meaning is possible to distinguish a topic from 
other objects in the context, so that the hearer will 
not necessarily guess the meaning used by the speaker. 
I will call this the gavagai-problem, because Quine 
used this word to illustrate exactly this difficulty.\footnote{See \cite{Quine:1960}, p. 29,30.}
Quine evoked the problem of an anthropologist
trying to figure out what a native speaking an unknown
language might mean when he utters `gavagai' while pointing
to a white rabbit scurrying by.

@@

Every time an agent successfully uses a form-meaning pair 
for speaking, he increments
the score with a specific increment $\delta$. 
$\delta$ is relatively high, typically equal to 0.1. 
The scores of competing associations, i.e. 
associations that used another form for the same meaning
are decremented with $\delta$. The score remains however bounded 
between 0.0 and 1.0. This way the ``best'' 
form-meaning pair stands out more clearly next time 
around.\footnote{A systematic investigation of alternatives for 
the updating function is contained in \cite{Oliphant:1997}. 
The dynamics of the mechanisms used in the 
Talking Heads experiments are being investigated in the
Ph.D thesis of Frederic Kaplan.}

@@

The Naming Game is a minimal model of communication 
between agents and far removed from the full complexity 
of human natural language. Moreover we made a number of 
simplifying assumptions, thus putting up scaffolds to
construct this initial model. The most important assumption
was that the meaning of a word can be unambiguously 
known by the speaker and hearer independently of language. 
This assumption is of course not valid for human 
beings, and neither is it valid for the Talking Heads.

>>>>

The \todo{This intro is too long} semiotic square captures the four entities in 
a linguistic interaction: 
the {\itshape referent}, which is the object in the physical world that the 
speaker wants to communicate, the {\itshape segmented image}, which is 
the internal perception of the referent, the {\itshape meaning}, which is a 
category or combination of categories that picks out the 
referent in the present context, and 
the {\itshape utterance}, which is the word form 
or set of word forms transmitted by the speaker. In the previous 
chapters we looked at two sides of the semiotic
square. The relation between the real world and the
perceived image was studied in chapter 3 and the relation
between the perceived image and a conceptualisation that 
could act as the meaning for a language communication was 
studied in chapter 4. We now turn to the next side of 
the semiotic square: the relation between meaning and 
utterance.

@@

Immediately we observe a major difficulty. In realistic
language games, where agents cannot inspect each others' 
brain states nor transmit meanings directly, 
there is no feedback about the 
meaning of a word, only about the referent. For example, 
when a speaker says `wabo' and the hearer has correctly pointed 
to the referent that the speaker intended, neither the 
speaker nor the hearer can know whether they were 
using the same meaning. They can only know that they 
arrived at the same referent. When a game fails, the speaker 
can only point to the topic and the hearer then 
tries to figure out what possible meaning 
could have been applicable. The speaker cannot
communicate directly the ``right'' meaning, and very often more 
than one meaning is possible to distinguish a topic from 
other objects in the context, so that the hearer will 
not necessarily guess the meaning used by the speaker. 
I will call this the gavagai-problem, because Quine 
used this word to illustrate exactly this difficulty.\footnote{See \cite[29--30]{Quine:1960}.}
Quine evoked the problem of an anthropologist
trying to figure out what a native speaking an unknown
language might mean when he utters `gavagai' while pointing
to a white rabbit scurrying by.

@@

Every time an agent successfully uses a form-meaning pair 
for speaking, he increments
the score with a specific increment $\delta$. 
$\delta$ is relatively high, typically equal to 0.1. 
The scores of competing associations, i.e. 
associations that used another form for the same meaning
are decremented with $\delta$. The score remains however bounded 
between 0.0 and 1.0. This way the ``best'' 
form-meaning pair stands out more clearly next time 
around.\todo{Cite Kaplan's thesis, please}\footnote{A systematic investigation of alternatives for 
the updating function is contained in \cite{Oliphant:1997}. 
The dynamics of the mechanisms used in the 
Talking Heads experiments are being investigated in the
Ph.D thesis of Frederic Kaplan.}

@@

\newpage
The Naming Game is a minimal model of communication 
between agents and far removed from the full complexity 
of human natural language. Moreover we made a number of 
simplifying assumptions, thus putting up scaffolds to
construct this initial model. The most important assumption
was that the meaning of a word can be unambiguously 
known by the speaker and hearer independently of language. 
This assumption is of course not valid for human 
beings, and neither is it valid for the Talking Heads.
